Geometric models reveal behavioral and neural signatures of transforming naturalistic experiences into episodic memories

### 2020-09-04 paper/main.tex
a couple more updates to results

\begin{figure}[tp]
  \centering
  \vspace*{-1cm}
  \includegraphics[width=.7\textwidth]{figs/precision_distinctiveness_detail}
  \caption{\small \textbf{Precision reflects the completeness of recall, whereas distinctiveness reflects recall specificity.} \textbf{A.} Recall precision by episode event.  Grey violin plots display kernel density estimates for the distribution of recall precision scores for a single episode event.  Colored dots within each violin plot represent individual participants' recall precision for the given event.  \textbf{B.} Recall distinctiveness by episode event. analogous to Panel A.  \textbf{C.} The set of ``Narrative Details'' episode annotations (generated by \citealp{ChenEtal17}) comprising an example episode event (22) identified by the HMM.  Each action or feature is highlighted in a different color.  \textbf{D.} Sentences comprising the most precise (P17) and least precise (P6) participants' recalls of episode event 22.  Descriptions of specific actions or features reflecting those highlighted in Panel B are highlighted in the corresponding color.  The text highlighted in gray denotes a (rare) false recall.  The unhighlighted text denotes correctly recalled information about other episode events.  \textbf{E.} The sets of ``Narrative Details'' episode annotations (generated by \citealp{ChenEtal17}) for scenes comprising episode events described by the example participants in Panel F.  Each event's text is highlighted in a different color.  \textbf{F.} The sentences comprising the most distinctive (P9) and least distinctive (P6) participants' recalls of episode event 22.  Sections of recall describing each each episode event in Panel E are highlighted with the corresponding color.}
  \label{fig:precision-detail}
\end{figure}

Examining individual recalls of the same episode event can provide insights into how the above precision and distinctiveness scores may be used to characterize similarities and differences in how different people describe the same shared experience.  In Figure \ref{fig:precision-detail}, we compare recalls for the same episode event from the participants with the highest (P17) and lowest (P6) precision scores.  From the HMM-identified episode event boundaries, we recovered the set of annotations describing the content of a single episode event (event 22; Fig.~\ref{fig:precision-detail}C), and divided them into different color-coded sections for each action or feature described.  Next, we used an analogous approach to identify the set of sentences comprising the corresponding recall event from each of the two example participants (Fig.~\ref{fig:precision-detail}D).  We then colored all words describing actions and features in the transcripts shown in Panel D according to the color-coded annotations in Panel C.  Visual comparison of these example recalls reveals that the more precise recall captures more of the episode event's content, and in greater detail.

Figure \ref{fig:precision-detail} also similarly illustrates the differences between high and low distinctiveness scores.  We extracted the set of sentences comprising the most distinctive recall event (P9) and least distinctive recall event (P6) corresponding to the example episode event shown in Panel C (event 22).  We also extracted the annotations for all episode events whose content these participants' single recall events described .  We assigned each episode event a unique color (Fig.~\ref{fig:precision-detail}E), and colored each recall sentence (Panel F) according to the episode events they best matched.  Visual inspection of Panel F reveals that the most distinctive recall's content is tightly concentrated around event 22, whereas the least distinctive recall incorporates content from a much wider range of episode events.

@@

The results thus far inform us about which aspects of the dynamic content in the episode participants watched were preserved or altered in participants' memories.  We next carried out a series of analyses aimed at understanding which brain structures might facilitate these preservations and transformations between the participants' shared experience of watching the episode and their subsequent memories of the episode.  In the first analysis, we sought to identify brain structures that were sensitive to the dynamic unfolding of the episode's content, as characterized by its topic trajectory.  We used a searchlight procedure to identify clusters of voxels whose activity patterns displayed a proximal temporal correlation structure (as participants watched the episode) matching that of the original episode's topic proportions (Fig.~\ref{fig:brainz}A; see \textit{Methods} for additional details).  In a second analysis, we sought to identify brain structures whose responses (during episode viewing) reflected how each participant would later structure their \textit{recounting} of the episode.  We used a searchlight procedure to identify clusters of voxels whose proximal temporal correlation matrices matched that of the topic proportions matrix for each participant's recall transcript (Figs.~\ref{fig:brainz}B; see \textit{Methods} for additional details).  To ensure our searchlight procedure identified regions \textit{specifically} sensitive to the temporal structure of the episode or recalls (i.e., rather than those with a temporal autocorrelation length similar to that of the episode and recalls), we performed a phase shift-based permutation correction (see \textit{Methods} for additional details). As shown in Figure~\ref{fig:brainz}C, the episode-driven searchlight analysis revealed a distributed network of regions that may play a role in processing information relevant to the narrative structure of the episode.  Similarly, the recall-driven searchlight analysis revealed a second network of regions (Fig.~\ref{fig:brainz}D) that may facilitate a person-specific transformation of one's experience into memory.  The top ten \texttt{Neurosynth} terms~\citep{YarkEtal11} associated with each (unthresholded) map are displayed in each panel.  In identifying regions whose responses to ongoing experiences reflect how those experiences will be remembered later, this latter analysis extends classic \textit{subsequent memory effect analyses}~\citep[e.g.,][]{PallWagn02} to the domain of naturalistic experiences.

@@

The searchlight analyses described above yielded two distributed networks of brain regions whose activity timecourses tracked with the temporal structure of the episode (Fig.~\ref{fig:brainz}C) or participants' subsequent recalls (Fig.~\ref{fig:brainz}D).  We next sought to gain greater insight into the structures and functional networks our results reflected.  To accomplish this, we performed an additional, exploratory analysis using \texttt{Neurosynth} \citep{YarkEtal11}.  Given an arbitrary statistical map as input, \texttt{Neurosynth} performs a massive automated meta-analysis, returning a ranked list of terms reported in papers with similar significance maps. We ran \texttt{Neurosynth} on the significance maps for the episode- and recall-driven searchlight analyses. These maps, along with the 10 terms with maximally similar meta-analysis images identified by \texttt{Neurosynth} are shown in Figure \ref{fig:brainz}.

>>>>

\begin{figure}[tp]
  \centering
  \vspace*{-1cm}
  \includegraphics[width=.7\textwidth]{figs/precision_distinctiveness_detail}
  \caption{\small \textbf{Precision reflects the completeness of recall, whereas distinctiveness reflects recall specificity.} \textbf{A.} Recall precision by episode event.  Grey violin plots display kernel density estimates for the distribution of recall precision scores for a single episode event.  Colored dots within each violin plot represent individual participants' recall precision for the given event.  \textbf{B.} Recall distinctiveness by episode event. analogous to Panel A.  \textbf{C.} The set of ``Narrative Details'' episode annotations (generated by \citealp{ChenEtal17}) comprising an example episode event (22) identified by the HMM.  Each action or feature is highlighted in a different color.  \textbf{D.} Sentences comprising the most precise (P17) and least precise (P6) participants' recalls of episode event 21.  Descriptions of specific actions or features reflecting those highlighted in Panel B are highlighted in the corresponding color.  The text highlighted in gray denotes a (rare) false recall.  The unhighlighted text denotes correctly recalled information about other episode events.  \textbf{E.} The sets of ``Narrative Details'' episode annotations (generated by \citealp{ChenEtal17}) for scenes comprising episode events described by the example participants in Panel F.  Each event's text is highlighted in a different color.  \textbf{F.} The sentences comprising the most distinctive (P9) and least distinctive (P6) participants' recalls of episode event 21.  Sections of recall describing each each episode event in Panel E are highlighted with the corresponding color.}
  \label{fig:precision-detail}
\end{figure}

Examining individual recalls of the same episode event can provide insights into how the above precision and distinctiveness scores may be used to characterize similarities and differences in how different people describe the same shared experience.  In Figure \ref{fig:precision-detail}, we compare recalls for the same episode event from the participants with the highest (P17) and lowest (P6) precision scores.  From the HMM-identified episode event boundaries, we recovered the set of annotations describing the content of a single episode event (event 21; Fig.~\ref{fig:precision-detail}C), and divided them into different color-coded sections for each action or feature described.  Next, we used an analogous approach to identify the set of sentences comprising the corresponding recall event from each of the two example participants (Fig.~\ref{fig:precision-detail}D).  We then colored all words describing actions and features in the transcripts shown in Panel D according to the color-coded annotations in Panel C.  Visual comparison of these example recalls reveals that the more precise recall captures more of the episode event's content, and in greater detail.

Figure \ref{fig:precision-detail} also similarly illustrates the differences between high and low distinctiveness scores.  We extracted the set of sentences comprising the most distinctive recall event (P9) and least distinctive recall event (P6) corresponding to the example episode event shown in Panel C (event 21).  We also extracted the annotations for all episode events whose content these participants' single recall events described .  We assigned each episode event a unique color (Fig.~\ref{fig:precision-detail}E), and colored each recall sentence (Panel F) according to the episode events they best matched.  Visual inspection of Panel F reveals that the most distinctive recall's content is tightly concentrated around event 21, whereas the least distinctive recall incorporates content from a much wider range of episode events.

@@

The results thus far inform us about which aspects of the dynamic content in the episode participants watched were preserved or altered in participants' memories.  We next carried out a series of analyses aimed at understanding which brain structures might facilitate these preservations and transformations between the participants' shared experience of watching the episode and their subsequent memories of the episode.  In the first analysis, we sought to identify brain structures that were sensitive to the dynamic unfolding of the episode's content, as characterized by its topic trajectory.  We used a searchlight procedure to identify clusters of voxels whose activity patterns displayed a proximal temporal correlation structure (as participants watched the episode) matching that of the original episode's topic proportions (Fig.~\ref{fig:brainz}A; see \textit{Methods} for additional details).  In a second analysis, we sought to identify brain structures whose responses (during episode viewing) reflected how each participant would later structure their \textit{recounting} of the episode.  We used a searchlight procedure to identify clusters of voxels whose proximal temporal correlation matrices matched that of the topic proportions matrix for each participant's recall transcript (Figs.~\ref{fig:brainz}B; see \textit{Methods} for additional details).  To ensure our searchlight procedure identified regions \textit{specifically} sensitive to the temporal structure of the episode or recalls (i.e., rather than those with a temporal autocorrelation length similar to that of the episode and recalls), we performed a phase shift-based permutation correction (see \textit{Methods} for additional details). As shown in Figure~\ref{fig:brainz}C, the episode-driven searchlight analysis revealed a distributed network of regions that may play a role in processing information relevant to the narrative structure of the episode.  Similarly, the recall-driven searchlight analysis revealed a second network of regions (Fig.~\ref{fig:brainz}D) that may facilitate a person-specific transformation of one's experience into memory.  In identifying regions whose responses to ongoing experiences reflect how those experiences will be remembered later, this latter analysis extends classic \textit{subsequent memory effect analyses}~\citep[e.g.,][]{PallWagn02} to the domain of naturalistic experiences.

@@

The searchlight analyses described above yielded two distributed networks of brain regions whose activity timecourses tracked with the temporal structure of the episode (Fig.~\ref{fig:brainz}C) or participants' subsequent recalls (Fig.~\ref{fig:brainz}D).  We next sought to gain greater insight into the structures and functional networks our results reflected.  To accomplish this, we performed an additional, exploratory analysis using \texttt{Neurosynth} \citep{YarkEtal11}.  Given an arbitrary statistical map as input, \texttt{Neurosynth} performs a massive automated meta-analysis, returning a ranked list of terms frequently used in neuroimaging papers that report similar statistical maps. We ran \texttt{Neurosynth} on the (unthresholded) permutation-corrected maps for the episode- and recall-driven searchlight analyses. The top ten terms with maximally similar meta-analysis images identified by \texttt{Neurosynth} are shown in Figure \ref{fig:brainz}.
